### 2015-02-16 paper/exprs.tex
expression typing

\begin{conj}[L-Value Use Preservation]
  If $\vdash\ltr$, $\tb{\ltr}{\gU}{\gG}{L}$, and $\uselv{\gG}{\gU}{\lv{x}{p}}{\gt}{\gU'}$
  then $\tb{\ltr}{\gU'}{\gG}{L}$.
\end{conj}

@@

\subsection*{Expression Typing}

>>>>

\begin{conj}[L-Value Use Shadow Preservation]
  If $\vdash\ltr$, $\tb{\ltr}{\gU}{\gG}{L}$, and $\uselv{\gG}{\gU}{\lv{x}{p}}{\gt}{\gU'}$
  then $\tb{\ltr}{\gU'}{\gG}{L}$.
\end{conj}

@@

\subsection*{Expression Typing}
With our model of the borrow checker in hand, we can finally specify the typing of expressions.
\newline

\fbox{$\tc{\gG;\ltr;L;\gU}{e}{\gt;\gU'}$}
\newline

Integer constants are trivial and do not change the shadow heap.

\begin{mathpar}
  \infer[ET-INT]{ }{\tc{\gG;\ltr;L;\gU}{z}{\tyint;\gU}}
\end{mathpar}

Using the value at a L-value simply relies on our earlier judgment.

\begin{mathpar}
  \infer[ET-USE]
    {\uselv{\gG}{\gU}{\lv{x}{p}}{\gt}{\gU'}}
    {\tc{\gG;\ltr;L;\gU}{\lv{x}{p}}{\gt;\gU'}}
\end{mathpar}

Creating an immutable borrowed reference requires checking that the immutable loan
guarantees will be upheld for the entire duration of the loan.
This involves checking that we can read the L-value now, and that we can guarantee
its validity and freezability for the duration of the loan.
We record the new loan in the output shadow heap.

\begin{mathpar}
  \infer[ET-REFIMM]
    {\tc{\gG}{\lv{x}{p}}{\gt} \\
     \canread{p}{\gU(x)} \\
     \valid{\gG}{\ltr}{L}{\lv{x}{p}}{\lt} \\
     \freezable{\gG}{\ltr}{\lv{x}{p}}{\lt} \\
     \textsc{record}(\gU,\lv{x}{p},(\lt,\qimm),\gU')
    }
    {\tc{\gG;\ltr;L;\gU}{\tyref{\lt}{\qimm}{\lv{x}{p}}}{\gt;\gU'}}
\end{mathpar}

Similarly, creating a mutable borrowed reference requires checking the mutable loan
guarantees. Instead of reading, we must be able to write. Instead of freezability,
we require uniqueness.

\begin{mathpar}
  \infer[ET-REFMUT]
  {\tc{\gG}{\lv{x}{p}}{\gt} \\
   \cw{p}{\gU(x)} \\
   \valid{\gG}{\ltr}{L}{\lv{x}{p}}{\lt} \\
   \unique{\gG}{\ltr}{\lv{x}{p}}{\lt} \\
   \textsc{record}(\gU,\lv{x}{p},(\lt,\qmut),\gU')
  }
  {\tc{\gG;\ltr;L;\gU}{\tyref{\lt}{\qmut}{\lv{x}{p}}}{\gt;\gU'}}
\end{mathpar}

\begin{conj}[Expression Shadow Preservation]
  If $\vdash\ltr$, $\tb{\ltr}{\gU}{\gG}{L}$, and $\tc{\gG;\ltr;L;\gU}{e}{\gt;\gU'}$
  then $\tb{\ltr}{\gU'}{\gG}{L}$.
\end{conj}

\begin{proof}[Justification]
  This is trivial in the integer case and follows from L-Value Use Shadow Preservation
  in the L-value use case, so we only need to discuss the borrow creation cases.
  In both borrow creation cases, the only change to the shadow heap is the addition of
  one new loan.

  When creating an immutable borrow, we know from \textsc{can-read} that
  no mutable loan exists on way to the path we are borrowing or in memory reachable from it.
  The shadow of the interior of the loaned path is now controlled by the immutable loan
  of the path, but since there were no mutable loans there before and it has not changed
  it should still be well-formed. Similarly, the shadows on the way to the path have no changed
  directly and they should not be affected by the addition of an immutable loan in their
  interior. The main concern is whether the addition of the new immutable loan to the bank 
  will result in a well-formed bank. We know from the validity check that the lifetime
  of the loan is within the lifetime bound of the bank. If the bank is empty or
  the lifetime of the old loan is not shorter than the lifetime of the new loan, then
  the bank will still be well-formed. We could restrict \textsc{record} to only succeed
  in this situation.

  When creating a mutable borrow, we know from \textsc{can-write} that
  no loan exists on the way to the path or in memory reachable from it.
  So similar to the immutable case, the shadow of the interior of the path will still
  be well-formed even though it is now controlled by a mutable loan.
  The shadows on the way to the path are still well-formed since they have no loans to begin with.
  As before, the only question is whether the bank of the path will remain well-formed.
  We know from \textsc{can-write} that the bank is empty, so the new bank, consisting of
  just one mutable loan, will be well-formed.
\end{proof}
